---
title: "RCG North-Sea and Eastern Arctic 2017:\\
	RDB exploration"
author: "RCG members"
date: "`r date()`"
output: 
  word_document
---

#Summary

A summary of the scope of this document.

#Practical information

This document is based on a Rmarkdown file. Tables, figures and numbers are
generated during the compilation of the original files using R code.
etc etc

#General options

In this section general option and parameters are set up :

```{r global_options, include=T,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
#knitr option
operationnel<-TRUE
knitr::opts_chunk$set(echo=FALSE, 
		      warning=!operationnel, 
		      message=!operationnel,
		      fig.height=6,
		      progress=!operationnel,
		      verbose=!operationnel,
		      include=TRUE,dev='png',autodep=FALSE)
#package
library(ggplot2)
library(dplyr)
library(DATRAS) #used to compute ices rectangle position

#path: user should set up the working path here if needed
#setwd("mypath")

#load some extra lookup table or data if needed
#load ices division definition (icesdiv.rdata is in the github depository)
load(file="icesdiv.rdata")
```

#Data

Here the data are read.

```{r dataaaaarg, include=T,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
 CL<-read.table(file="../../../data/CL 2009-2016 NSEA.csv",sep=",",header=T)
 CE<-read.table(file="../../../data/CE 2009-2016 NSEA.csv",sep=",",header=T)
 #TR<-read.table(file="../../../data/TR NSEA.csv",sep=",",header=T)
 #HH<-read.table(file="../../../data/HH NSEA.csv",sep=",",header=T)
 #CA<-read.table(file="../../../data/CA NSEA.csv",sep=",",header=T)
 #2016 only
 CL<-CL%>%filter(Year==2016)
 CE<-CE%>%filter(Year==2016)
```

#Landings

Landings by countries and more.

```{r landings_barplot, include=T,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
	#summarize information
 	ts1<-tbl_df(CL)%>%
		group_by(LandingCountry)%>%
		summarise(w=sum(OfficialLandingCatchWeight,na.rm=T)/1000)%>%
		arrange(w)
	#plot information
	ts1plt<-ggplot(ts1,aes(x=LandingCountry,y=w))+
		geom_bar(stat="identity")+
		xlab("Landing Country")+ylab("Landings (t)")+
		ggtitle(paste0("Landings by countries"))
	print(ts1plt)
```

```{r landings_map, include=T,cache=F,echo=F,warning=F,message=F,progress=F,verbose=F}
	#summarise data: year/trim
	map1<-CL%>%group_by(Area)%>%summarise(w=sum(OfficialLandingCatchWeight,na.rm=T)/1000)%>%ungroup()
	icesdivtmp<-inner_join(icesdiv,map1,by=c("area"="Area"))
	iddivtmp<-iddiv%>%filter(area%in%unique(icesdivtmp$area))%>%group_by(area)%>%summarise(x=mean(x),y=mean(y))%>%
		ungroup()%>%mutate(shortarea=gsub("\\.","",gsub("27.","",area)))
	map1<-left_join(map1,iddivtmp,by=c("Area"="area"))
	map1moy<-map1%>%group_by(Area,x,y)%>%summarise(w=mean(w))
	map2<-CL%>%group_by(StatisticalRectangle)%>%summarise(w=sum(OfficialLandingCatchWeight,na.rm=T)/1000)%>%ungroup()
	map2$lon<-map2$lat<-map2$lonc<-map2$latc<-NA
	for(i in 1:nrow(map2)){
		if(nchar(map2$StatisticalRectangle[i])==4 & map2$StatisticalRectangle[i]!="NULL"){
		map2$lon[i]<-DATRAS::icesSquare2coord(map2$StatisticalRectangle[i],"midpoint")$lon
		map2$lat[i]<-DATRAS::icesSquare2coord(map2$StatisticalRectangle[i],"midpoint")$lat
		map2$lonc[i]<-DATRAS::icesSquare2coord(map2$StatisticalRectangle[i],"corner")$lon
		map2$latc[i]<-DATRAS::icesSquare2coord(map2$StatisticalRectangle[i],"corner")$lat
		}
	}
	rangex<-c(min(map2$lon,na.rm=T)-.5,max(map2$lon,na.rm=T)+.5)
	rangey<-c(min(map2$lat,na.rm=T)-.5,max(map2$lat,na.rm=T)+.5)

	#poly map
		map<-ggplot()+theme_bw()+
			theme(panel.grid.minor.y= element_blank(),
		              panel.grid.minor.x = element_blank())+
			#geom_polygon(data=icesdivtmp,aes(x=long,y=lat,group=group),fill=NA,border="black")+
			geom_path(data=icesdivtmp,aes(x=long,y=lat,group=group),color="dark grey")+#,fill=NA,border="black")+
			geom_raster(data=map2,aes(x=lon,y=lat,fill=w),stat="identity",alpha=.75)+ 
			scale_fill_distiller(palette='Spectral',name="Landings (t)")+
			geom_point(data=map1,aes(x=x,y=y,size=w),alpha=.25)+
			geom_point(data=map1moy,aes(x=x,y=y,size=w),alpha=.5,col="red",fill=NA,shape=1)+
			scale_size_continuous(name="Landings (t, by area)")+
			#geom_polygon(data=coast_map,aes(x=long,y=lat,group=group),fill="grey")+#coord_fixed(1)
			#geom_vline(xintercept=seq(-12, 30, by=1),col="light grey")+
			#geom_hline(yintercept=seq(39,90, by=0.5),col="light grey")+
			borders("worldHires",xlim=rangex,ylim=rangey,fill="light grey",colour="light grey")+
			geom_text(data=iddivtmp,aes(x=x,y=y+1.5,label=shortarea))+
			coord_quickmap(xlim=range(icesdivtmp$long),ylim=range(icesdivtmp$lat))+
			xlab("")+ylab("")+
			ggtitle(paste0("Landings"))
```

#Efforts

Effort by countries and more.

```
```


